---
title: "UE20CS312 - Data Analytics - Worksheet 1a - Part 2 - Exploring data with R"
subtitle: "PES University"
author: 
  - 'VISMAYA.R, Dept. of CSE - PES2UG20CS391'
output: pdf_document
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

## PART I - Exploratory Data Analysis with R

### Loading dataset and inserting libraries

```{r}
library(tidyverse)
library(ggplot2)
cbc_df <- read_csv("CharlesBookClubDataset.csv")
```

### Problem 1

Generate an understanding of the dataset via a summary of its features.
Find the count, missing count, minimum, 1st quartile, median, mean, 3rd
quartile, max and standard deviation of all relevant columns.
Separately, print the total number of missing values in each column.

```{r}
summary(cbc_df)
# Number of missing values in each column
colSums(is.na(cbc_df))
```

### Problem 2

Replace missing values within the Recency, Frequency, and Monetary
features with suitable values. Explain your reasoning behind the method
of substitution used.. Also, plot the distribution of feature values
after imputation.

```{r}
#plotting histogram before replacing missing values
hist(cbc_df$M, breaks=30,main="Monetary before filling missing values")
hist(cbc_df$R, breaks=30, main="Recency before filling missing values")
hist(cbc_df$F, breaks=30, main="Frequency before filling missing values")

mode <- function(x) {
mode0 <- names(which.max(table(x)))
    if(is.numeric(x)) return(as.numeric(mode0))
    return(mode0)
}
# Apply to all necessary columns
cbc_df$R[is.na(cbc_df$R)] <- mode(cbc_df$R)
cbc_df$F[is.na(cbc_df$F)] <- mode(cbc_df$F)
cbc_df$M[is.na(cbc_df$M)] <- mode(cbc_df$M)

```

By the histograms, we get to know that the 3 columns M,R and F are
positively skewed , hence the method chosen for data imputaion will be
Replacing by mode. The distribution of feature values after data
imputation is as follows:

```{r}
#plotting histogram after replacing missing values
hist(cbc_df$'M', main="Monetary after filling missing values", breaks=30)
hist(cbc_df$'R', main="Recency after filling missing values", breaks=30)
hist(cbc_df$'F', main="Frequency after filling missing values", breaks=30)
```

### Problem 3

Discretize the continuous values of Monetary, Recency, and Frequency
into appropriate bins, and create three new columns Mcode, Rcode and
Fcode respectively, for the discretized values. Print out the summary of
the newly created columns.Explicitly mention the number of bins used and
explain the choice for the bin size. Print out the summary of the newly
created columns.What are the most optimum breakpoints you can choose?
Try to think of a statistical function that provides these breakpoints
for optimum binning.

```{r}
#problem 3

cbc_df <- cbc_df %>% mutate(Rcode=cut(cbc_df$R,
  breaks=unique(
  quantile(cbc_df$R,
  probs=seq.int(0,1,by=1/4))),
  include.lowest=TRUE),
  
Mcode=cut(cbc_df$M,
  breaks=unique(
  quantile(cbc_df$M,
  probs=seq.int(0,1,by=1/5))),
  include.lowest=TRUE),

  Fcode=cut(cbc_df$F,
  breaks=unique(
  quantile(cbc_df$F,
  probs=seq.int(0,1,by=1/4))),
  include.lowest=TRUE))

levels(cbc_df$Mcode) <- c('$15-$112', '$112-$181', '$181-$242', '$242-$296', '$296-$479')
levels(cbc_df$Rcode) <- c('2-8 months', '8-14 months', '14-16 months', '16-36 months')
levels(cbc_df$Fcode) <- c('1-2 books', '2-6 books', '6-12 books')

cat("\n")
#summary
print("Mcode:")
summary(cbc_df$Mcode)
cat("\n")
print("Rcode:")
summary(cbc_df$Rcode)
cat("\n")
print("Fcode:")
summary(cbc_df$Fcode)

```

### Problem 4

#### Problem 4.1

Transform the Florence variable into a categorical feature that can take
up the values True or False. Create and visualize horizontal box plots
for the original Recency, Frequence, Monetary and FirstPurch features
against the Florence variable.

```{r}
#Problem 4
#Problem 4.1
ggplot(cbc_df, aes(x = Fcode)) + geom_bar()+labs(x = "Frequency")
ggplot(cbc_df, aes(x = Rcode)) + geom_bar()+labs(x = "Recency")
ggplot(cbc_df, aes(x = Mcode)) + geom_bar()+labs(x = "Monetary")
ggplot(cbc_df, aes(x = FirstPurch)) + geom_bar()+labs(x = "First Purchase")
```

#### Problem 4.2

Transform the Florence variable into a categorical feature that can take
up the values True or False. Create and visualize horizontal box plots
for the original Recency, Frequence, Monetary and FirstPurch features
against the Florence variable.

```{r}
#Problem 4.2
#Transforming "Florence"

cbc_df$Florence <- factor(cbc_df$Florence, labels = c("No", "Yes"))
# Plot box plots
ggplot(cbc_df, aes_string(x = "Florence", y = "R")) +
geom_boxplot() +
coord_flip() +
labs(x = "Recency", y = "Did the customer make a purchase?") 

ggplot(cbc_df, aes_string(x = "Florence", y = "M")) +
geom_boxplot() +
coord_flip()+
labs(x = "Recency", y = "Did the customer make a purchase?") 

ggplot(cbc_df, aes_string(x = "Florence", y = "F")) +
geom_boxplot() +
coord_flip() +
labs(x = "Frequency", y = "Did the customer make a purchase?")
```

#### Problem 4.3

Create and visualize a density plot for Recency, Frequency, Monetary and
FirstPurch features.

```{r}
#Problem 4.3
ggplot(cbc_df, aes_string(x = "R", fill = "Florence")) +
geom_density(alpha = 1) +
labs(x = "Recency") +
theme(legend.position = c(0.75, 0.75))

ggplot(cbc_df, aes_string(x = "M", fill = "Florence")) +
geom_density(alpha = 1) +
labs(x = "Monetary") +
theme(legend.position = c(0.9, 0.9))

ggplot(cbc_df, aes_string(x = "F", fill = "Florence")) +
geom_density(alpha = 1) +
labs(x = "Frequency") +
theme(legend.position = c(0.5, 0.5))
```

## Part II. ANOVA

### Problem 1

1.  Considering the dataset,we use a one-way ANOVA test. As the data
    contains one categorical independent variable and one quantitative
    dependent variable. Also, the independent variable has more than
    three levels.

2.  We can perform an ANOVA in R using the functions: aov() ,anova() ,

```{r}
library(ggpubr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(broom)
library(car)
df <- read.csv('Scenario 1.csv') #Load the Dataset

anova_1 <-aov(df$No.of.items ~ df$POI, data = df)
summary(anova_1)
```

3.The Df column displays that the degrees of freedom for the POI
variable is 4, and the degrees of freedom for the residuals (the total
number of observations minus one and minus the number of levels in the
POI variable) is 995

The Sum Sq column displays the sum of squares, this means that the total
variation between the group means and the overall mean is 127 for the
POI variable and for the Residuals it is, 30827.

The Mean Sq column is the mean of the sum of squares, calculated to be
31.75 for the POI column and 30.98 for the Residuals.

The F-value column is the test statistic from the F test.

The Pr(\>F) column is the p-value of the F-statistic.The p-value of the
POI variable is greater than the significance level (p \> 0.05), so it
appears that the POI does not have a real impact on the No.of.items in
the evidence log about him/her.

### Problem 2

1.  Considering the dataset,we use a two-way ANOVA test. As the data
    contains more than one categorical independent variable and one
    quantitative dependent variable. Also, the independent variables
    have more than three levels.
2.  We can perform an ANOVA in R using the functions: aov() ,anova() ,

```{r}
library(ggpubr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(broom)
library(car)
df2 <- read.csv('Scenario 2.csv') #Load the Dataset

anova_2 <-aov(df2$No.of.items ~ df2$POI * df2$Priority, data = df2)
summary(anova_2)
```

3.The Df column displays that the degrees of freedom for the POI
variable.

The Sum Sq column displays the sum of squares.

The Mean Sq column is the mean of the sum of squares.

The F-value column is the test statistic from the F test.

Pr(\>F) is the p-value of the F-statistic. Considering 5% significance,
it is clear that both POI and Priority columns have a relation with the
No of items in the evidence log. Also, as both POI and Priority columns
are categorical, there can be no relation between them.

4.assumptions of 2 way ANOVA are:

i)Independence of observations: Categorical variables cannot be compared
with F Statistic and can only be ensured to be independent variables by
experimental design.

ii) Normally-distribution of dependent variable

iii) From the following graphs Homogeneity of variance is clear.

```{r}
plot(anova_2)

hist(df2$No.of.items,main="Dependent Variable",xlab="No of Items")
hist(anova_2$residuals,main="Histogram of Residuals",xlab="Residuals")

```

### Problem 3

```{r}
library(ggpubr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(broom)
library(car)
df2 <- read.csv('Scenario 2.csv') #Load the Dataset

anova_2 <-aov(formula=df2$No.of.items ~ factor(df2$POI) + df2$Priority, data = df2)

tukey_anova_2<-TukeyHSD(anova_2)

tukey_anova_2

par(mar=c(5,8,4,1)+.1)
tukey.plot.test<-TukeyHSD(aov(formula = No.of.items ~ factor(POI), data = df2))
plot(tukey.plot.test, las = 1)

par(mar=c(5,8,4,1)+.1)
tukey.plot.test<-TukeyHSD(aov(formula = No.of.items ~ Priority, data = df2))
plot(tukey.plot.test, las = 1)

```

From the post-hoc test results, we see that there are no statistically
significant differences (p \< 0.05) between the different sets of POI
groups. There is no difference in the average no of Evidence items
collected when compared with any two POI.

While, there is a statistically significant difference between the
Priority groups: high-critical,low-critical and med-critical. The other
Priority groups show no statistically significant difference.This shows
that assigning high priority to cases does bring about change in the
collection of evidence.
