---
title: "UE20CS312 - Data Analytics - Worksheet 1b"
subtitle: "PES University"
author: 
  - 'VISMAYA.R, Dept. of CSE - PES2UG20CS391'
output: pdf_document
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

### Loading dataset and inserting libraries

```{r}
library(tidyverse)
library(ggpubr)
library(ggplot2)
df <- read.csv('road_accidents_india_2016.csv', row.names=1)
```

### Problem 1

```{r}
Total_acc<-function(wcol,tcol){
  
  for(i in 1:36) {
   tcol[i] <- tcol[i] + wcol[i] 
  }
  return(tcol)
}

total_col <- seq(1,36)
for(j in 1:36){total_col[j]<-0}
df$total_accidents <-Total_acc(df$Fine.Clear...Total.Accidents,total_col)
df$total_accidents <-Total_acc(df$Mist..Foggy...Total.Accidents,df$total_accidents)
df$total_accidents <-Total_acc(df$Cloudy...Total.Accidents,df$total_accidents)
df$total_accidents <-Total_acc(df$Rainy...Total.Accidents,df$total_accidents)

df$total_accidents <-Total_acc(df$Snowfall...Total.Accidents,df$total_accidents)
df$total_accidents <-Total_acc(df$Hail.Sleet...Total.Accidents,df$total_accidents)
df$total_accidents <-Total_acc(df$Dust.Storm...Total.Accidents,df$total_accidents)
df$total_accidents <-Total_acc(df$Others...Total.Accidents,df$total_accidents)

for(k in 1:36){cat("Total accidents in",df$State..UT[k],":",df$total_accidents[k],fill=TRUE)}
```

### Problem 2

```{r}
#Finding fatality rate
total_col <- seq(1,36)
for(j in 1:36){total_col[j]<-0}
df$Total_deaths<- Total_acc(df$Fine.Clear...Persons.Killed,total_col)
df$Total_deaths<- Total_acc(df$Mist..Foggy...Persons.Killed,df$Total_deaths)
df$Total_deaths<- Total_acc(df$Cloudy...Persons.Killed,df$Total_deaths)
df$Total_deaths<- Total_acc(df$Rainy...Persons.Killed,df$Total_deaths)

df$Total_deaths<- Total_acc(df$Snowfall...Persons.Killed,df$Total_deaths)
df$Total_deaths<- Total_acc(df$Hail.Sleet...Persons.Killed,df$Total_deaths)
df$Total_deaths<- Total_acc(df$Dust.Storm...Persons.Killed,df$Total_deaths)
df$Total_deaths<- Total_acc(df$Others...Persons.Killed,df$Total_deaths)

#for(k in 1:36){cat("Total deaths in",df$State..UT[k],":",df$Total_deaths[k],fill=TRUE)}

frate<-function(tacc,tdeath){
  fcol <- seq(1,36)
  for(i in 1:36) {
    fcol[i] <- tdeath[i] / tacc[i] 
  }
  return(fcol)
}

df$fatality_rate <-frate(df$total_accidents,df$Total_deaths)
for(k in 1:36){cat("Fatality rate in",df$State..UT[k],":",df$fatality_rate[k],fill=TRUE)}
```

Finding Mist/Foggy rate:

```{r}
#Mist/foggy rate:Fraction of total accidents that happen in mist/foggy conditions
mfgrate<-function(tacc,mfgacc){
  mfgcol <- seq(1,36)
  for(i in 1:36) {
    mfgcol[i] <- mfgacc[i] / tacc[i] 
  }
  return(mfgcol)
}
df$mfg_rate <-mfgrate(df$total_accidents,df$Mist..Foggy...Total.Accidents)
for(k in 1:36){cat("Mist/Foggy rate in",df$State..UT[k],":",df$mfg_rate[k],fill=TRUE)}
```

## Finding correlation coefficent

```{r}
result=cor.test(df$fatality_rate,df$mfg_rate , method = "pearson")
print(result)
```

Since the p-value is greater than 0.05 (5% significance level), there is
no statistically significant correlation between the fatality rate and
the mist/foggy rate

## Scatter PLot

```{r}
plot(df$fatality_rate,df$mfg_rate,cex=2, col="#69b3a2",xlab="Fatality rate", ylab="misty/foggy rate",main="Plot of fatality rate vs mist/foggy rate" )
```

## PROBELM 3

```{r}
df$acc_rank<-rank(desc(df$total_accidents),ties.method='random')
df$fatal_rank <- rank(desc(df$Total_deaths),ties.method='random')

print(cor.test(df$acc_rank,df$fatal_rank, method='spearman'))

```

```{r}
ggscatter(df, x='acc_rank', y='fatal_rank',
add='reg.line', conf.int=TRUE,
cor.coef=TRUE, cor.method = 'spearman')
```

```{r}
rs <- cor(df$acc_rank,df$fatal_rank,
method='spearman')

degrees <- nrow(df) - 2
t_stat <- (rs - 0.2)/sqrt((1 - rs*rs)/(nrow(df) - 2))
t_stat

#2-tailed test p-value
2 * pt(q=t_stat, df=degrees, lower.tail=FALSE)
```

Since the p-value is lesser than 0.05 (5% significance level),our
hypotheses is correct

## Problem 4

```{r}
df$hail_binary <- ifelse(df$Hail.Sleet...Total.Accidents > 0, 1, 0)
df$rain_acc <- df$Rainy...Total.Accidents
print.data.frame(df[, c('State..UT','hail_binary', 'rain_acc')])

```

```{r}
cor.test(df$rain_acc,df$hail_binary, method='pearson')
```

There is no significant correlation

## Probelm 5

```{r}
df$dust_binary <- ifelse(df$Dust.Storm...Total.Accidents > 0, 1, 0)
contingency_table <- table(df[, c('dust_binary', 'hail_binary')])
contingency_table

library(psych)
phi(contingency_table)

```
